---
title: "R Markdown for datasets cleaning"
author: ""
date: "2025-03-14"
output: html_document
---



## Data Cleaning

After the successful scraping of data, the data science team wants to ensure that the data is cleaned and prepared for analysis, in order to ensure that it is accurate, consistent, and reliable. 

The team will deal with missing data values and outliers. Data types would be standardized, and the data structure would be made to be consistent across the entire dataset. 

The data would be handled in a methodical and reproducible format to maintain integrity for subsequent analysis.


```{r}
install.packages("tidyverse")
install.packages("mice")
```



```{r}
#loading of the required library for data cleaning
library(tidyverse)
library(mice)

```


```{r}

#loading of Toyota dataframe into R environment
getwd()
#setwd("C:/Users/HP/Desktop/Data mining pre datasets")
toyota <- read.csv ("toyota.csv")

head(toyota)

```

```{r}

# to separate 'Car_Brand' column into 'Brand' and 'Model' columns so we can get the brand name and their models separately. This is for Toyota df.
toyota_cleaned <- toyota %>%
  separate(
    Car_Brand, into = c("Brand", "Model"), 
    sep = " ", 
    extra = "merge", 
    fill = "right"
  )

head(toyota_cleaned)

```


```{r}

# to clean 'Price' by removing non-numeric characters and converting to numeric
toyota_cleaned <- toyota_cleaned %>%
  #modifies or add a column in a dataframe
  mutate(
    Price = as.numeric(gsub("[^0-9.]", "", Price))  #gsub function replaces anything that is not 0-9 with an empty string. and then, they are converted to numeric.
  )

# Rename 'Price' column for clarity
colnames(toyota_cleaned)[colnames(toyota_cleaned) == "Price"] <- "Price_in_Pounds"

# After converting to numeric, check for NA values
sum(is.na(toyota_cleaned$Price_in_Pounds))



head(toyota_cleaned)

```



```{r}

# Clean 'Mileage' by removing non-numeric characters and converting to numeric
toyota_cleaned <- toyota_cleaned %>%
  mutate(
    Mileage = as.numeric(gsub("[^0-9.]", "", Mileage))
  )

# Rename 'Mileage' column for clarity
colnames(toyota_cleaned)[colnames(toyota_cleaned) == "Mileage"] <- "Mileage_in_Miles"

# After converting to numeric, check for NA values
sum(is.na(toyota_cleaned$Mileage_in_Miles))

# Rename 'release_year' column for clarity
colnames(toyota_cleaned)[colnames(toyota_cleaned) == "release_year"] <- "Release_Year"


head(toyota_cleaned)

```


```{r}

# Check for null values in each column
null <- colSums(is.na(toyota_cleaned))
print("Null Values per Column:")
print(null)

# View dataset structure
str(toyota_cleaned)

#write.csv(toyota_cleaned, "toyota_cleaned.csv", row.names = FALSE)

```

```{r}
#load the ford data into the environment
getwd()
#setwd("C:/Users/HP/Desktop/Data mining pre datasets")
ford <- read.csv ("ford.csv")

head(ford)

```


```{r}
# to separate 'Car_Brand' column into 'Brand' and 'Model' columns so we can get the brand name and their models separately. This is for Ford df.
ford_cleaned <- ford %>%
  mutate(
    # Split Brand and Model based on position of the alphabets. 
    Model = substr(Brand, 5, nchar(Brand)), #extract from the 5th character to the end and  in brand and send into column model
    Brand = substr(Brand, 1, 4)  #extract the first 4 characters from brand and send to column Brand
  )

head(ford_cleaned)  # Display the first few rows

  
```

```{r}

# Rename 'mileage' column to 'Mileage'
colnames(ford_cleaned)[colnames(ford_cleaned) == "mileage"] <- "Mileage"


# Now convert 'Price' and 'Mileage' to numeric, removing none numeric character
ford_cleaned = ford_cleaned %>%
  mutate(
    Price = as.numeric(gsub("[^0-9]", "", Price)),
    Mileage = as.numeric(gsub("[^0-9]", "", Mileage))  
  )

head(ford_cleaned)

```


```{r}

# Reorder 'ford_cleaned' column names
ford_cleaned = ford_cleaned %>%
  select(Brand, Model, Price, Transmission, fuel_Type, Mileage, release_year) %>%
  
  # Rename the columns for consistency
  rename(
    "Price_in_Pounds" = Price,
    "Mileage_in_Miles" = Mileage,
    "Fuel_Type" = fuel_Type,
    "Release_Year" = release_year
  )

head(ford_cleaned)


```


```{r}

# Check for null values in the cleaned dataset
null <- colSums(is.na(ford_cleaned))
print("Null Values per Column:")
print(null)

# View dataset structure
str(ford_cleaned)


#write.csv(ford_cleaned, "ford_cleaned.csv", row.names = FALSE)


```



```{r}
#load the bmw data into R environment 
getwd()
#setwd("C:/Users/HP/Desktop/Data mining pre datasets")
bmw <- read.csv("bmw.csv")

head(bmw)

```



```{r}

# Prevent scientific notation
options(scipen=999)

# Data Cleaning & Transformation
bmw_cleaned <- bmw %>%
  rename(Model = Brand) %>%  # Rename existing Brand column to Model
  mutate(
    Brand = "BMW",  # Add a new Brand column with "BMW" for all rows
    Mileage = na_if(Mileage, ""),  # Convert empty strings to NA
    Mileage = na_if(Mileage, " "), # Convert space-only cells to NA
    Price = as.numeric(gsub("[^0-9]", "", Price)),  # Remove non-numeric characters
    Mileage = as.numeric(gsub("[^0-9]", "", Mileage)),
    FuelType = tolower(FuelType)  # Convert FuelType to lowercase
  ) %>%
  rename(
    Price_in_Pounds = Price,
    Mileage_in_Miles = Mileage,
    Fuel_Type = FuelType,
    Release_Year = release_year
  ) %>%
  select(Brand, everything())  # Move the new Brand column to the first position

# Display first few rows
head(bmw_cleaned)


```



```{r}

# Impute missing values for Mileage using the median per Model, rounding the values
bmw_cleaned <- bmw_cleaned %>%
  group_by(Model) %>%
  mutate(Mileage_in_Miles = ifelse(is.na(Mileage_in_Miles), 
                                   round(median(Mileage_in_Miles, na.rm = TRUE)), 
                                   round(Mileage_in_Miles))) %>%
  ungroup()

# Display first few rows
head(bmw_cleaned)


```



```{r}

# Check for remaining missing values after imputation
print("Null Values after Imputation:")
print(colSums(is.na(bmw_cleaned)))

# View dataset structure
str(bmw_cleaned)

# Save cleaned dataset
#write.csv(bmw_cleaned, "bmw_cleaned.csv", row.names = FALSE)

```


```{r}
#checking for outliers for toyota price and mileage
boxplot(toyota_cleaned$Price_in_Pounds, toyota_cleaned$Mileage_in_Miles, 
        main = "Outliers for Toyota Price and Mileage",
        names = c("Price in Pounds", "Mileage in Miles"),
        col = c("lightblue", "lightgreen"),
        outline = TRUE  # Keep outliers visible
        )

```



```{r}
#checking for outliers for FORD price and mileage
boxplot(ford_cleaned$Price_in_Pounds, ford_cleaned$Mileage_in_Miles, 
        main = "Outliers for Ford Price and Mileage",
        names = c("Price in Pounds", "Mileage in Miles"),
        col = c("lightblue", "lightgreen"),
        outline = TRUE  # Keep outliers visible
        )
```


```{r}


#checking for outliers for bmw miles and mileage
boxplot(bmw_cleaned$Price_in_Pounds, bmw_cleaned$Mileage_in_Miles, 
        main = "Outliers for BMW Price and Mileage",
        names = c("Price in Pounds", "Mileage in Miles"),
        col = c("lightblue", "lightgreen"),
        outline = TRUE  # Keep outliers visible
        )
```


```{r}
#replacing outliers in price and mileage with NAs for toyota
 for (x in c('Price_in_Pounds', 'Mileage_in_Miles'))
{
  value = toyota_cleaned[,x][toyota_cleaned[,x] %in% boxplot.stats(toyota_cleaned[,x])$out]
  toyota_cleaned[,x][toyota_cleaned[,x] %in% value] = NA
}

# Check the number of missing values AFTER removing outliers
print("Missing values AFTER replacing outliers with NA:")
print(colSums(is.na(toyota_cleaned)))
```



```{r}

# Replacing outliers in price and mileage with NA for BMW
for (x in c("Price_in_Pounds", "Mileage_in_Miles")) {
  outliers <- boxplot.stats(bmw_cleaned[[x]])$out  # Identify outliers
  bmw_cleaned[[x]][bmw_cleaned[[x]] %in% outliers] <- NA  # Replace with NA
}

# Check the number of missing values AFTER removing outliers
print("Missing values AFTER replacing outliers with NA:")
print(colSums(is.na(bmw_cleaned)))


```

```{r}

#performing imputation using mice for toyota
impute_toyota_cleaned <- mice(toyota_cleaned)
completed_toyota_cleaned <- complete(impute_toyota_cleaned, 1)

head(completed_toyota_cleaned)
as.data.frame(colSums(is.na(completed_toyota_cleaned)))

```


```{r}

#performing imputation using mice for bmw
impute_bmw_cleaned <- mice(bmw_cleaned)
completed_bmw_cleaned <- complete(impute_bmw_cleaned, 1)

head(completed_bmw_cleaned)
as.data.frame(colSums(is.na(completed_bmw_cleaned)))

```


```{r}

#checking for outliers for toyota after imputation
boxplot(completed_toyota_cleaned$Price_in_Pounds, completed_toyota_cleaned$Mileage_in_Miles, 
        main = "Outliers for Toyota Price and Mileage after Imputation",
        names = c("Price in Pounds", "Mileage in Miles"),
        col = c("lightblue", "lightgreen"),
        outline = TRUE  # Keep outliers visible
        )

```



```{r}

#checking for outliers for bmw after imputation
boxplot(completed_bmw_cleaned$Price_in_Pounds, completed_bmw_cleaned$Mileage_in_Miles, 
        main = "Outliers for BMW Price and Mileage after Imputation",
        names = c("Price in Pounds", "Mileage in Miles"),
        col = c("lightblue", "lightgreen"),
        outline = TRUE  # Keep outliers visible
        )

```


```{r}

#replacing remaining outliers after imputation with NAs for toyota
 for (x in c('Price_in_Pounds', 'Mileage_in_Miles'))
{
  value = completed_toyota_cleaned[,x][completed_toyota_cleaned[,x] %in% boxplot.stats(completed_toyota_cleaned[,x])$out]
  completed_toyota_cleaned[,x][completed_toyota_cleaned[,x] %in% value] = NA
}

# Check the number of missing values AFTER removing outliers
print("Missing values AFTER replacing outliers with NA:")
print(colSums(is.na(completed_toyota_cleaned)))

```



```{r}

#replacing remaining outliers after imputation with NAs for bmw
 for (x in c('Mileage_in_Miles'))
{
  value = completed_bmw_cleaned[,x][completed_bmw_cleaned[,x] %in% boxplot.stats(completed_bmw_cleaned[,x])$out]
  completed_bmw_cleaned[,x][completed_bmw_cleaned[,x] %in% value] = NA
}

# Check the number of missing values AFTER removing outliers
print("Missing values AFTER replacing outliers with NA:")
print(colSums(is.na(completed_bmw_cleaned)))

```

```{r}

# drop NAs after imputation, and Check for remaining missing values
completed_toyota_cleaned <- drop_na(completed_toyota_cleaned, c(Price_in_Pounds, Mileage_in_Miles))
as.data.frame(colSums(is.na(completed_toyota_cleaned)))


completed_bmw_cleaned <- drop_na(completed_bmw_cleaned, c(Mileage_in_Miles))
as.data.frame(colSums(is.na(completed_bmw_cleaned)))

```


```{r}

# Merge all three datasets into one
merged_data <- bind_rows(completed_toyota_cleaned, ford_cleaned, completed_bmw_cleaned)

# Check the first few rows of the merged dataset
head(merged_data)

#write.csv(merged_data, "merged_data.csv", row.names = FALSE)

```


```{r}

# Check for missing values in each column
missing_values <- colSums(is.na(merged_data))
print("Missing values per column:")
print(missing_values)

# Final structure of the dataset
print("Structure of the merged dataset:")
str(merged_data)


```

