---
title: "Data Mining Portfolio2 web scraping"
author: "Falcon_Group"
group_members: 
Deborah Bose Adebayo  - 2401695
Daniel Femi Babalola  - 2418023
Akinpelu Akintomide   - 2408921
Oluyinka Abosede Adekoya - 2409914
output: html_document
date: "2025-03-17"
---



## Business Problem

A leading car retail company with branches across the uk, name The Company is looking to form used car dealership partnership with the top five most popular brands in the country. According to Statista (2022), these brands are Volkswagen, Audi, Ford, BMW and Toyota. However, only three of these brands (Ford, Toyota and BMW) have shown interest in the partnership.

Before going into any agreement, The Company seeks to leverage consumer market intelligence as a guide to strike a deal that aligns with consumer attitudes, and profitability.To achieve this, the data science department has decided to scrape data on Car Brand, Model, Year, Millage, Transmission and Fuel type from the website of the three brands.

Before scrapping, the data science team carefully goes through each website policy to check existing regulations that restricts the scraping exercise, they found none. Also the regulations were sent to the company's legal department for extra confirmation.


```{r}
#Installing of necessary packages for the successful parsing html content from a website. The two packages to be used are Rselenium and rvest, other installed packages are for a smooth running of the scraping process. 

install.packages("RSelenium") # For automating web browsers for web scraping
install.packages("rvest") # For scraping data from HTML web pages
install.packages("dplyr") 
install.packages("tidyverse")
install.packages("netstat")#used to access free port 
install.packages("wdman") #Web driver Manager for 
install.packages("stringr") #extracting string characters
install.packages("binman") #to check Chrome version for web scraping

```


```{r}

#Loading of the installed packages
library(RSelenium)
library(rvest)
library(dplyr)
library(tidyverse)
library(netstat)
library(wdman)
library(stringr)
library(binman)

```


```{r}
#download the Rselenium web driver  
selenium()
# This checks for the location where our drivers are stored. check=F to disable check notifications
selenium_obj <- selenium(retcommand = T, check = F) 
```

```{r}

# Check if the Chrome driver is available and its version
binman::list_versions("chromedriver")

```


```{r}
#setting up the Rselenuim server for chrome browser
rs_driver_obj1 <- rsDriver(browser = "chrome", 
                          chromever = "134.0.6998.88",
                          verbose = FALSE)
```

```{r}
#opening the client side of or remote browser, which controls the browser
rD <- rs_driver_obj1$client 
rD$open()
```

```{r}
#An empty dataframe to hold the scraped data in the loop 
dataloop <- data.frame() 

# Looping through pages 1 to 21 of the toyota used car page
for (i in 1:2) {
  # Construct URL for each page
  looplink <- paste0("https://used.toyota.co.uk/approved-used?step=carFilter&imagery[]=10829&mileage=8000&deposit=250&term=48&monthlypay=425&payinfull=1000,200000&depositMultiple=6&maintenance=0&financeGroupType=personal&acceptedFinance=false&p=", i)
  
  # Navigate to the webpage
  rD$navigate(looplink)
  
  # Allow time for page load
  Sys.sleep(3)
  
  # Accept cookie policy if button exists. The tryCatch function allows the loop to continue running if the agree button is not found, popping out the error message.
  tryCatch({
    agree_button <- rD$findElement(using = "css selector", value = "#onetrust-accept-btn-handler") #Access the agree button using find element.
    agree_button$clickElement()
    Sys.sleep(2) # Give time after clicking
  }, error = function(e) { message("No cookie popup found, continuing...") })
  
  # Get page source from object rD which houses the client side of our remote browser using Rslenium functions. 
  page_source <- rD$getPageSource()[[1]]
  page1 <- read_html(page_source %>% unlist()) #Read and save  HTML content to page1 object/variable
  
  # Extract data from the HTML content page1 using "rvest" function html_nodes.
  Brand <- html_nodes(page1, ".title") %>%
    html_text(trim = TRUE) %>%
    .[2:13]  # Ensure correct range
  
  Price <- html_nodes(page1, ".price-wrapper") %>%
    html_text(trim = TRUE) %>% #removes white space before or after the html texts
    str_remove_all("Cash Purchase|\\n")  # Removes unwanted text
  
  Transmission <- html_nodes(page1, ".pod-data") %>%
    html_text() %>%
    str_extract("\\b(Manual|Automatic)\\b") %>% #only extract either of the both option where available 
    str_trim()
  
  FuelType <- html_nodes(page1, ".pod-data") %>%
    html_text() %>%
    str_extract("\\b(Diesel|Electric|Petrol|Petrol Hybrid|Petrol Plug-in Hybrid)\\b") %>% #only extract either of the both option where available(stringr)
    str_trim() 
  
  Mileage <- html_nodes(page1, ".pod-data") %>%
    html_text(trim = TRUE) %>%
    str_extract("\\d{1,3},\\d{3}\\sMiles") %>% # Extract numbers from thousand to hundreds that appear before miles. extract miles with it. stringr function is used here
    str_trim()
  
  release_year <-  html_nodes(page1, ".pod-data") %>%
    html_text() %>% 
    str_extract("\\d{4}") # extracts only thousands number where available 
  
  # Combine into dataframe
  toyota_test_df <- data.frame(
    Car_Brand = Brand,
    Price = Price,
    Transmission = Transmission,
    Fuel_Type = FuelType,
    Mileage = Mileage,
    release_year = release_year,
    stringsAsFactors = FALSE
  )
  
  # Append to dataloop
  dataloop <- bind_rows(dataloop, toyota_test_df)
  
  # Sleep before next iteration
  Sys.sleep(3)
}
view(dataloop)

write_csv(dataloop, "testtoyota.csv", na = "")


```

```{r}
dataloop1 <- data.frame()


#dataloop <- data.frame()

# Loop through pages 1 to 20 of the ford usedcar pages
for (i in 1:2) {
  # Construct URL for each page
  looplink1 <- paste0("https://secure.ford.co.uk/shop/price-and-locate/approved-used/direct-used-cars/results#/searchResults?distance=9999&locationId=ChIJee4lya2ne0gRJwAcmXLAkZM&longLatCoordinates=-2.4359159,53.5732043&page=",i)
  
  # Navigate to the webpage
  rD$navigate(looplink1)
  
  # Allow time for page load
  Sys.sleep(3)
  
  # Accept cookie policy if button exists
  tryCatch({
    agree_button1 <- rD$findElement(using = "css selector", value = "a.initialized:nth-child(2)")
    agree_button1$clickElement()
    Sys.sleep(2) # Give time after clicking
  }, error = function(e) { message("No cookie popup found, continuing...") })
  
  # Get page source
  page_source1 <- rD$getPageSource()[[1]]
  page2 <- read_html(page_source1 %>% unlist())
  
  # Extract data
  Brand <- html_nodes(page2, ".model-title") %>%
    html_text() %>% str_replace_all("[\n\t]", "") %>%
    str_trim()
  
  price <- html_nodes(page2, ".model-attributes-item") %>%
    html_text() %>% str_extract("£\\d{1,3},\\d{3}") %>%
    na.omit() %>% .[1:12] 
  
  Transmission_type <- html_nodes(page2, ".model-attributes-item") %>%
    html_text() %>%
    str_extract("\\b(Automatic|Manual|PowerShift)\\b")%>%
    na.omit() %>% .[1:12]
  
  fuel_type <- html_nodes(page2, ".model-attributes-item") %>%
    html_text() %>%
    str_extract("\\b(Diesel|Electric|Petrol|LPG|CNG|Bi-Fuel|Flexi-Fuel|Hybrid|Plug-in Hybrid)\\b")%>%
    na.omit() %>% .[1:12]
  
  mileage <- html_nodes(page2, ".model-attributes-item") %>%
    html_text() %>%
    str_extract("\\d{1,3},\\d{3}\\sMiles|\\d{3}\\sMiles")%>% 
    na.omit() %>% .[1:12]
  
  release_year2 <- html_nodes(page2, ".model-attributes-item") %>% 
    html_text() %>% 
    str_extract("\\d{4}") %>% 
    na.omit() %>% .[1:12]
  
  
  # Combine into dataframe
  ford_df <- data.frame(
    Brand = Brand,
    Price = Price,
    Transmission = Transmission,
    fuel_Type = FuelType,
    mileage = Mileage,
    release_year = release_year2,
    stringsAsFactors = FALSE
  )
  
  # Append to dataloop
  dataloop1 <- bind_rows(dataloop1, ford_df)
  
  # Sleep before next iteration
  Sys.sleep(3)
}
view(dataloop1)

write_csv(dataloop1, "testford.csv", na = "")

```

```{r}
dataloop2 <- data.frame()

# Loop through pages 1 to 11 of the BMW used car pages
for (i in 1:2) {
  # Construct URL for each page
  looplink2 <- paste0("https://usedcars.bmw.co.uk/result/?page=", i, "&series=BMWi&size=23")
  
  # Navigate to the webpage
  rD$navigate(looplink2)
  
  # Allow time for page load
  Sys.sleep(3)
  
  # Accept cookie policy if button exists
  tryCatch({
    accept_button2 <- rD$findElement(using = "css selector", value = ".accept-button")
    accept_button2$clickElement()
    Sys.sleep(2) # Give time after clicking
  }, error = function(e) { message("No cookie popup found, continuing...") })
  
  # Get page source
  page_source2 <- rD$getPageSource()[[1]]
  page2 <- read_html(page_source2 %>% unlist())
  
  # Extract data
  Brand <- html_nodes(page2, ".uvl-c-overview-card__title") %>% html_text()
  
  Price <- html_nodes(page2, ".uvl-c-quote-tile__cash") %>% 
    html_text() %>% 
    str_extract("\\d{1,3},\\d{3}")
  
  Transmission <- html_nodes(page2, ".uvl-c-overview-card-specs") %>% 
    html_text() %>% 
    str_extract("(?i)Automatic|Manual")
  
  FuelType <- html_nodes(page2, ".uvl-c-overview-card-specs") %>% 
    html_text() %>% 
    str_extract("(?i)ELECTRIC|HYBRID")
  
  Mileage <- html_nodes(page2, ".uvl-c-overview-card-specs") %>% 
    html_text() %>% 
    str_extract("(?i)(\\d{1,3},\\d{3}|\\d{1,3})\\s*miles")
  
  release_year3 <- html_nodes(page2, ".uvl-c-overview-card-specs") %>%
    html_text() %>% 
    str_extract("\\d{4}")
  
  # Combine into dataframe
  bmw_df <- data.frame(
    Brand = Brand,
    Price = Price,
    Transmission = Transmission,
    FuelType = FuelType,
    Mileage = Mileage,
    release_year = release_year3,
    stringsAsFactors = FALSE
  )
  
  # Append to dataloop2
  dataloop2 <- bind_rows(dataloop2, bmw_df)
  
  # Sleep before next iteration
  Sys.sleep(3)
}
view(dataloop2)

write_csv(dataloop2, "testbmw.csv", na = "")
```


```{r}
#stop the Rselenium server from running.
rD$close()
```

